Reconstructing Propositional Proofs in Type Theory

### 2017-10-23 paper/sections/proof-reconstruction/resolve.tex
[ paper ] fixed rm funnctions and example added.

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal forms,
and reordering those and the inner disjunctions to match them.
Below, we provide some lemmas to perform such reordering tasks,
we omit the proofs but they can found in~\cite{AgdaMetis}.

% \begin{itemize}
%   \item Reorder a disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$ to match with
% $ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$. For all $i$, there exists at least one $j$ such
% that $φᵢ ≡ ψⱼ$.
%   \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
% CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$. For all $i$, there exists at least one $j$
% such that $φᵢ ≡ ψⱼ$.
% \end{itemize}

First, we define \fassoc$_{⬓}$ function in~(\ref{def:right-square})
to convert a disjunction or a conjunction into its right-associative form.
The square symbol can be the conjunction symbol or the disjunction symbol.
We use $\fassoc_{∧}$ to convert conjunctions and $\fassoc_{∨}$ for
disjunctions.

\begin{equation}
\label{def:right-square}
  \begin{aligned}
    &\hspace{.495mm}\fassoc_{⬓} : \Prop \to \Prop\\
    &\begin{array}{lll}
    \fassoc_{⬓} &((φ₁~⬓~φ₂)~⬓~φ₃) &= \fassoc_{⬓}~(φ₁~⬓~(φ₂~⬓~φ₃))\\
    \fassoc_{⬓} &(φ₁~⬓~φ₂)        &= φ₁~∨~\fassoc_{⬓}~φ₂\\
    \fassoc_{⬓} &φ                &= φ
    \end{array}
  \end{aligned}
\end{equation}

@@

\begin{mainlemma}
\label{lem:build-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fbuild_{∨}~φ~ψ$, where,
\begin{equation*}
  \begin{split}
  \label{eq:build-or}
  &\fbuild_{∨} : \Prop → \Prop → \Prop\\
  &\fbuild_{∨}~φ~ψ~ =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } ψᵢ ≡ \fbuild_{∨}~φ~ψᵢ\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation*}
\end{mainlemma}

From now, we assume all propositions to be right-associative unless
otherwise stated.
The $\ffactor$ function in~(\ref{eq:factor-definition})
simplifies a special case of disjunction, the repeated disjuncts
(\eg, $\ffactor(φ ∨ φ) = φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $ψ ∨ φ$.

>>>>

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal form,
and reordering those and the inner disjunctions to match them.
Below, we provide some lemmas to perform such reordering tasks,
we omit the proofs but they can found in~\cite{AgdaMetis}.

First, we define \fassoc$_{⬓}$ function in~(\ref{eq:fassoc})
to convert a disjunction or a conjunction into its right-associative form.
The square symbol can be the conjunction symbol or the disjunction symbol.
We use $\fassoc_{∧}$ to convert conjunctions and $\fassoc_{∨}$ for
disjunctions.

\begin{equation}
\label{eq:fassoc}
  \begin{aligned}
    &\hspace{.495mm}\fassoc_{⬓} : \Prop \to \Prop\\
    &\begin{array}{lll}
    \fassoc_{⬓} &((φ₁~⬓~φ₂)~⬓~φ₃) &= \fassoc_{⬓}~(φ₁~⬓~(φ₂~⬓~φ₃))\\
    \fassoc_{⬓} &(φ₁~⬓~φ₂)        &= φ₁~∨~\fassoc_{⬓}~φ₂\\
    \fassoc_{⬓} &φ                &= φ
    \end{array}
  \end{aligned}
\end{equation}

@@

\begin{mainlemma}
\label{lem:build-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fbuild_{∨}~φ~ψ$, where,
\begin{equation}
  \begin{split}
  \label{eq:build-or}
  &\fbuild_{∨} : \Prop → \Prop → \Prop\\
  &\fbuild_{∨}~φ~ψ~ =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } ψᵢ ≡ \fbuild_{∨}~φ~ψᵢ\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation}
\end{mainlemma}

From now, we assume all propositions to be right-associative unless
otherwise stated.
The $\ffactor$ function in~(\ref{eq:factor-definition})
simplifies a special case of disjunction, the repeated disjuncts
(\eg, $\ffactor~(φ ∨ φ)~=~φ$).
Notice that other cases like $φ~∨~(ψ~∨~φ)$ do no reduce to $ψ~∨~φ$.
